Optimal (Multiway) Spatial Joins

### 2024-05-23 msj.tex
yufei

\begin{figure*}
    \begin{tabular}{cccc}
        \includegraphics[height=30mm]{./artwork/prob-a} & 
        \hspace{3mm}
        \includegraphics[height=30mm]{./artwork/prob-b} &
        \hspace{3mm}
        \includegraphics[height=30mm]{./artwork/prob-c} &
        \hspace{3mm}
        \includegraphics[height=30mm]{./artwork/prob-d} \\
        (a) Problem $\EuScript{A}$ &
        \hspace{3mm}
        (b) Problem $\EuScript{B}$ &
        \hspace{3mm}
        (c) Problem $\EuScript{C}$ &
        \hspace{3mm}
        (d) Problem $\EuScript{D}$
    \end{tabular}

@@

\extraspacing {\bf Problem $\bm{\EuScript{A}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that there exists at least one rectangle $r \in R$ covering both the left end point of $h$ and the point $(x, y)$. If the point $p$ exists (equivalently, $h$ has at least one left-end covering rectangle in $R$), then we should output a tuple $(h, p)$; otherwise, we output nothing for $h$. Figure~\ref{fig:probs}a gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.

@@

\extraspacing {\bf Problem $\bm{\EuScript{B}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each rectangle $r \in R$, its lowest crossing segment $h \in H$ --- formally, $h = [x_1, x_2] \times y$ has the smallest $y$ coordinate among all the segments in $H$ crossing $r$. If such an $h$ does not exist, output nothing for $r$.  Figure~\ref{fig:probs}b gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1, h_3), (r_2, h_2)}$.

@@

\extraspacing {\bf Problem $\bm{\EuScript{D}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each rectangle $r \in R$, all the segments $h \in H$ crossing $r$. Furthermore, those segments need to be sorted by their y-coordinates. Formally, if $h_1, h_2, ..., h_t$ (for some $t \ge 1$) are all the segments in $H$ crossing $r$, we output $(r: h_1, h_2, ..., h_t)$, provided that $y_i > y_{i-1}$ for each $i \in [2, t]$ where $y_i$ (resp., $y_{i-1}$) is the y-projection of $h_i$ (resp., $h_{i-1}$). Figure~\ref{fig:probs}d gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1: h_3, h_2), (r_2: h_2, h_1)}$.

@@

@@

A remark is in order about duplicate removal. A result tuple of Type 2 can belong to, in the worst case, all the $4(k-2)$ sub-types. Hence, by considering each sub-type separately, we may see the same result tuple multiple times (precisely, up to $4(k-2)$ times) in the whole algorithm. However, this does {\em not} mean that the tuple needs to be {\em reported} multiple times. Whenever a Type-2 result tuple is found, we can immediately decide in $O(k)$ time all the sub-types it belongs to. To avoid outputting the tuple more than once, we can enforce a policy to designate a specific sub-type for outputting. One such policy is the following: among all sub-types that the tuple belongs to, identify the one with the smallest sub-type number $t$ (an integer from 1 to $4(k-2)$); report the tuple only when we are computing the particular sub-type $t$.

@@

@@

>>>>

\begin{figure*}
    \begin{tabular}{ccc}
        \includegraphics[height=30mm]{./artwork/prob-a} & 
        \hspace{3mm}
        \includegraphics[height=30mm]{./artwork/prob-b} &
        \hspace{3mm}
        \includegraphics[height=30mm]{./artwork/prob-c} \\
%         &
%         \hspace{3mm}
%         \includegraphics[height=30mm]{./artwork/prob-d} \\
        (a) Problem $\EuScript{A}$ and $\EuScript{D}$ &
        \hspace{3mm}
        (b) Problem $\EuScript{B}$ &
        \hspace{3mm}
        (c) Problem $\EuScript{C}$
%         &
%         \hspace{3mm}
%         (d) Problem $\EuScript{D}$
    \end{tabular}

@@

\extraspacing {\bf Problem $\bm{\EuScript{A}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each rectangle $r \in R$, its lowest crossing segment $h \in H$ --- formally, $h = [x_1, x_2] \times y$ has the smallest $y$ coordinate among all the segments in $H$ crossing $r$. If such an $h$ does not exist, output nothing for $r$.  Figure~\ref{fig:probs}b gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1, h_3), (r_2, h_2)}$.

@@

\extraspacing {\bf Problem $\bm{\EuScript{B}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that there exists at least one rectangle $r \in R$ covering both the left end point of $h$ and the point $(x, y)$. If the point $p$ exists (equivalently, $h$ has at least one left-end covering rectangle in $R$), then we should output a tuple $(h, p)$; otherwise, we output nothing for $h$. Figure~\ref{fig:probs}a gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$.

@@

\extraspacing {\bf Problem $\bm{\EuScript{D}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each rectangle $r \in R$, all the segments $h \in H$ crossing $r$. Furthermore, those segments need to be sorted by their y-coordinates. Formally, if $h_1, h_2, ..., h_t$ (for some $t \ge 1$) are all the segments in $H$ crossing $r$, we output $(r: h_1, h_2, ..., h_t)$, provided that $y_i > y_{i-1}$ for each $i \in [2, t]$ where $y_i$ (resp., $y_{i-1}$) is the y-projection of $h_i$ (resp., $h_{i-1}$). See Figure~\ref{fig:probs}a again where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1: h_3, h_2), (r_2: h_2, h_1)}$.

@@

Recall that $n$ is the total size of $R_1, ..., R_{k-2}, H$, and $V$. Henceforth, we assume that a distinct ID from 1 to $n$ is given to each rectangle or segment in $R_1, ..., R_{k-2}, H$, and $V$.

@@

A remark is in order about duplicate removal. A result tuple of Type 2 can belong to, in the worst case, all the $4(k-2)$ sub-types. Hence, by considering each sub-type separately, we may see the same result tuple multiple times (precisely, up to $4(k-2)$ times) in the whole algorithm. However, this does {\em not} mean that the tuple needs to be {\em reported} multiple times. Whenever a type-2 result tuple is found, we can immediately decide in $O(k)$ time all the sub-types it belongs to. To avoid outputting the tuple more than once, we can enforce a policy to designate a specific sub-type for outputting. One such policy is the following: among all sub-types that the tuple belongs to, identify the one with the smallest sub-type number $t$ (an integer from 1 to $4(k-2)$); report the tuple only when we are computing the particular sub-type $t$.

@@

This subsection will explain how to find the $k$-tuples in $\J(R_1, ..., R_{k-2},$ $H, V)$ that belong to Type 1. We will represent the set of such tuples as $\J_1(R_1, ..., R_{k-2}, H, V)$ and define its ``projection'' onto $(R_1, ..., R_{k-2})$ as
\myeqn{
    && \Pi_{1:k-2}(\J_1(R_1, ..., R_{k-2}, H, V)) \nn \\
    &=& \{(r_1, ..., r_{k-2}) \mid (r_1, ..., r_{k-2}, h, v) \in \J_1(R_1, ..., R_{k-2}, H, V)
    \}. \nn
}
We will compute $\J_1(R_1, ..., R_{k-2}, H, V)$ in two steps:
\myitems{
    \item {{\bf Step 1:}} Compute $\Pi_{1:k-2}(\J_1(R_1, ..., R_{k-2}, H, V))$;
    \item {{\bf Step 2:}} Generate $\J_1(R_1, ..., R_{k-2}, H, V)$ from the result of Step 1.
}

\noindent {\bf Step 1.} We will accomplish this step by solving an instance of $(k-2)$-SJ constructed as follows. For each rectangle $r_1 \in R_1$, we compute four segments:
\myitems{
    \item $h_1$: the lowest segment (i.e., with the smallest y-coordinate) among all the segments in $H$ that cross $r$;
    \item $h_2$: the highest segment (i.e., with the largest y-coordinate) among all the segments in $H$ that cross $r$;
    \item $v_1$: the leftmost segment (i.e., with the smallest x-coordinate) among all the segments in $V$ that cross $r$;
    \item $v_2$: the rightmost segment (i.e., with the largest x-coordinate) among all the segments in $V$ that cross $r$.
}
Define $r_1'$ the rectangle defined by the above four segments in the way illustrated in Figure~\ref{fig:four-segs}, where $r_1'$ is the shaded area. Note that $r_1'$ exists if and only if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$. It is possible for $h_1$ and $h_2$ to be the same segment (same for $v_1$ and $v_2$).

\begin{figure}
    \includegraphics[height=25mm]{./artwork/four-segs}
    \figcapup
    \caption{Constructing $r_1'$ from $r_1$}
    \label{fig:four-segs}
    \figcapdown
\end{figure}

a rectangle $r_1'$ to an initially empty set $R_1'$ if $r_1$ is crossed by at least one horizontal segment in $H$ and by at least one vertical segment in $V$. Specifically,

@@

\section*{Appendix}
